Extracting Function Signature from R Program Traces

### 2020-05-13 paper/main.tex
data.frames

\caption{Struct example from the \code{neuralnet} package.}
\label{fig:general-validation-example}
\end{center}
\end{figure}

@@

%
%
\subsubsection{Forgoing Data Frames}

Speaking of classes, an important classful value in R is the data frame.
Data frames underpin nearly all idiomatic use of the language, and even
popular \code{tibble} (from the \code{tidyverse} ecosystem) and
\code{data.table} (from the \code{data.table} package) build on the
essential type.  We entertained the possibility of including data frame
types, but \AT{found them to be clunky? Question mark?}  Consider:
\AT{example of why data.frame types would be bad atm.}

Data frames do have a type, through the \code{class<data.frame>} type, and
future work will expand on this type, discussed further in
Section~\ref{sec:futurework}.

>>>>

\caption{Struct example from the \code{neuralnet} package.}
\label{fig:struct-ex}
\end{center}
\end{figure}

@@

%
%
\subsubsection{Data Frames}

One of the most popular classes in R is the \code{data.frame} class \AT{we have numbers here, grab em}.
Data frames and the derivative \code{tibble} (from the \code{tidyverse} ecosystem) and \code{data.table} (from the \code{data.table} package) types underpin a huge amount of data analysis in R.

One way to deal with data frames is through the struct type, with a named field for each column of the data frame, but as mentioned previously structs introduced undue noise into \typetracer's analysis results. 
Further complicated issues for data frames is that many functions built to operate on them operate in a name-agnostic way.
For instance, the \code{tidyverse} package ecosystem allows programmers to pass column names to functions which operate on their data frames.
In base R, idiomatic data frame use is to use string column names to select rows from the frame (unless only a single column is of interest, wherein the \code{$} syntax is appropriate). 
We include some examples in Figure~\ref{fig:data-frames-bad} to illustrate the complexities.

\begin{lstlisting}
# Let's put some examples showing that data frames can get out of hand, and that
# structs are inadequate.
\end{lstlisting}

\caption{Select examples of idiomatic \code{data.frame} use.}
\label{fig:data-frames-bad}
\end{center}
\end{figure}

Doing justice to this type will be an endeavour in and of itself, requiring reconciliation of base R's \code{data.frame}, \code{tidyverse}'s \code{tibble}, and \code{data.table}'s \code{data.table}.
\AT{We could talk about how metaprogramming is pervasive in dplyr, and how we would need to capture that, too, which is a huge undertaking.}
It will also require proper treatment of function types, and perhaps even parametric types, which dynamic analyses such as \typetracer are unable to discern soundly.
\AT{General words to convince the reader that data frames are, in fact, hard, and worthy of their own paper.}

In the context of our type annotation language, we will ascribe the type \code{class<data.frame>} to \code{data.frame}-typed values.
A more in-depth discussion of class types follows.
